%% sections/abstract.tex

\begin{abstract}
Dong drum towers are emblematic timber heritage structures of the Dong
ethnic group in Southwest China. Their layered eaves, dense timber
frameworks, roof-crown systems, and shadowed internal spaces pose
substantial challenges for comprehensive digital documentation. This
study establishes an open, conservation-oriented benchmark for UAV-based
neural reconstruction of four representative Dong drum towers---Chaoli,
Congjiang, Zeli, and Zengchong. Four neural reconstruction methods,
including Neural Radiance Fields (NeRF), 2D Gaussian Splatting (2DGS),
3D Gaussian Splatting (3DGS), and Instant-NGP (INGP), were evaluated
under a unified and reproducible workflow that integrates UAV image
acquisition, SfM-based camera pose estimation, train/test splitting,
scene-level reconstruction, and standardized evaluation. The benchmark
provides open access to the dataset, source code, experimental
parameters, train/test splits, and reproduction protocols through a
public GitHub repository. Rendering quality was assessed using PSNR,
SSIM, and LPIPS, together with training efficiency, rendering speed,
and component-level visual interpretation. The results show that 3DGS
achieved the strongest overall image fidelity and structural similarity,
with average PSNR and SSIM values of 31.60~dB and 0.924, respectively.
INGP obtained the best average LPIPS value of 0.075 and the shortest
training time, indicating its advantage for rapid reconstruction and
perceptual visualization. NeRF produced stable but less efficient
results, whereas 2DGS showed limited performance in image-level
rendering metrics under the current configuration, although its
surface-oriented formulation remains relevant for future
geometry-focused evaluation. By combining an open benchmark,
quantitative metrics, architectural component readability, and
conservation-oriented interpretation, this study provides a
reproducible and task-aware basis for selecting neural reconstruction
methods for complex timber heritage documentation.
\end{abstract}

\begin{IEEEkeywords}
Dong drum tower, neural reconstruction, UAV photogrammetry,
Neural Radiance Fields (NeRF), 3D Gaussian Splatting (3DGS),
2D Gaussian Splatting (2DGS), Instant-NGP, heritage documentation,
benchmark
\end{IEEEkeywords}
