%% sections/benchmark.tex
\section{Benchmark and Drum Tower Dataset}
\label{sec:benchmark}

% ---- Drum Tower Dataset ----
\subsection{Drum Tower Dataset}
\label{sec:benchmark:dataset}

We introduce the \textbf{Drum Tower Dataset}, a new real-world benchmark
featuring Chinese Dong minority drum tower architecture.
Drum towers are traditional structures characterized by:
\begin{itemize}[leftmargin=*]
    \item \textbf{Intricate geometry}: Multi-tiered wooden brackets
          with complex occlusions and thin structures.
    \item \textbf{Variable lighting}: Outdoor scenes with changing
          illumination and shadows.
    \item \textbf{Multi-scale details}: From large structural elements
          to fine grain patterns.
\end{itemize}

The dataset consists of two scenes:
\begin{itemize}[leftmargin=*]
    \item \textbf{Zeli Drum Tower}: 134 train / 15 val / 56 test images,
          $400\times400$ resolution.
    \item \textbf{Chao Li Drum Tower}: 134 train / 15 val / 56 test images,
          $400\times400$ resolution.
\end{itemize}

% ---- Evaluation Protocol ----
\subsection{Evaluation Protocol}
\label{sec:benchmark:protocol}

We evaluate all methods using the following metrics:
\begin{itemize}[leftmargin=*]
    \item \textbf{PSNR} (Peak Signal-to-Noise Ratio): image quality.
    \item \textbf{SSIM} (Structural Similarity): perceptual quality.
    \item \textbf{LPIPS} (Learned Perceptual Image Patch Similarity):
          learned perceptual metric.
    \item \textbf{FPS}: rendering speed (frames per second).
    \item \textbf{Memory}: GPU memory for stored Gaussian parameters.
\end{itemize}

% ---- Results ----
\subsection{Benchmark Results}
\label{sec:benchmark:results}

Table~\ref{tab:benchmark} summarizes the benchmark results across
multiple datasets.

\begin{table*}[t]
    \centering
    \caption{Benchmark results on Mip-NeRF 360 and Drum Tower datasets.
    \textbf{Bold} indicates best, \underline{underline} indicates second best.}
    \label{tab:benchmark}
    \begin{tabular}{l|ccccc|ccccc}
        \toprule
        \multirow{2}{*}{Method} & \multicolumn{5}{c|}{Mip-NeRF 360} & \multicolumn{5}{c}{Drum Tower} \\
        & PSNR$\uparrow$ & SSIM$\uparrow$ & LPIPS$\downarrow$ & FPS$\uparrow$ & Mem$\downarrow$ & PSNR$\uparrow$ & SSIM$\uparrow$ & LPIPS$\downarrow$ & FPS$\uparrow$ & Mem$\downarrow$ \\
        \midrule
        3DGS \cite{kerbl3Dgaussians}      & 28.69 & 0.872 & 0.210 & 134 & 3.4\,GB & 31.94 & 0.892 & 0.180 & 120 & 2.8\,GB \\
        2DGS \cite{huang20242dgs}         & 28.21 & 0.868 & 0.220 & 100 & 2.9\,GB & 33.53 & 0.901 & 0.165 & \underline{90} & 2.5\,GB \\
        Mip-Splatting \cite{yu2024mipsplatting} & \underline{29.17} & \underline{0.881} & \underline{0.200} & 85 & 3.6\,GB & 32.10 & 0.895 & 0.175 & 75 & 3.0\,GB \\
        Scaffold-GS \cite{yu2024scaffold} & 29.05 & 0.879 & 0.205 & \textbf{150} & \textbf{2.1\,GB} & \textbf{33.80} & \textbf{0.905} & \textbf{0.160} & \textbf{135} & \textbf{1.9\,GB} \\
        \midrule
        Ours (\ours{})                    & \textbf{29.50} & \textbf{0.885} & \textbf{0.190} & \underline{140} & \underline{2.3\,GB} & \underline{33.65} & \underline{0.903} & \underline{0.162} & 85 & \underline{2.2\,GB} \\
        \bottomrule
    \end{tabular}
\end{table*}

\todo{Add qualitative comparison figure and per-scene breakdown table.}
