%% sections/conclusion.tex
\section{Conclusion}
\label{sec:conclusion}

\section{Limitations and Future
Work}\label{limitations-and-future-work}

This study has several limitations. First, the evaluation is primarily
image-based. Without independent high-precision reference geometry,
metric accuracy indicators such as cloud-to-cloud distance,
cloud-to-mesh distance, Chamfer distance, Hausdorff distance, and
completeness were not computed. Second, the corpus includes only four
representative scenes; broader validation requires more tower types,
regional variations, roof-tier densities, settlement contexts, and
acquisition conditions. Third, the Chaoli scene lacks an independent
LLFF-style test set, so its quantitative values should be treated as
reference values rather than fully comparable test-set results. Fourth,
GCP/RTK information and weather or lighting metadata were not available
for all scenes, which may affect scale reliability and cross-scene
comparability. Finally, downstream conservation tasks such as component
segmentation, HBIM object generation, deformation monitoring, damage
comparison, and multi-temporal assessment were not evaluated.

Future work should introduce independent geometric references, expand
the open corpus, standardize acquisition metadata, and test how
reconstruction quality affects HBIM modeling, component recognition,
measurement reliability, and conservation-oriented interpretation.

\section{Conclusion}\label{conclusion}

This study established an open, conservation-oriented benchmark for
UAV-based neural reconstruction of Dong drum towers. Four
scenes---Chaoli, Congjiang, Zeli, and Zengchong---were used to compare
NeRF, 2DGS, 3DGS, and INGP under a unified and reproducible workflow.
The released benchmark includes the dataset, code, parameters,
train/test splits, and reproduction protocols.

3DGS achieved the strongest image fidelity and structural similarity,
making it the most suitable evaluated method for high-fidelity visual
documentation and structurally readable reconstruction. INGP achieved
the best average LPIPS and the shortest training time, supporting rapid
reconstruction and preliminary visualization. NeRF remained a stable but
less efficient baseline. 2DGS underperformed under image-level metrics,
but its surface-oriented formulation remains relevant for
geometry-focused assessment.

The main conclusion is that method selection for timber heritage
reconstruction should be task-oriented. 3DGS is better suited to
high-fidelity documentation and visual interpretation, while INGP is
more practical for rapid and lightweight reconstruction.
Geometry-sensitive tasks such as measurement, HBIM object extraction,
deformation monitoring, and surface-based documentation require further
validation with independent reference data. By linking open UAV heritage
data, neural reconstruction methods, standardized evaluation, and
conservation-oriented interpretation, this study provides a reproducible
basis for method selection in complex timber heritage documentation.

\textbf{Data Availability Statement}

The dataset, source code, experimental parameters, train/test
splits, and reproduction protocols generated in this study are openly
available in the project GitHub repository.
