%% sections/introduction.tex
\section{Introduction}
\label{sec:introduction}

% ---- Motivation ----
Novel view synthesis (NVS) has long been a central problem in computer
vision and graphics. The introduction of Neural Radiance Fields (NeRF)
\cite{mildenhall2020nerf} revolutionized the field by leveraging
coordinate-based neural networks to encode scene radiance.
However, NeRF-based methods suffer from slow training and rendering
speeds due to expensive volumetric ray-marching.

3D Gaussian Splatting (3DGS) \cite{kerbl3Dgaussians} addressed these
limitations by representing scenes as collections of anisotropic 3D
Gaussians, rendered via a differentiable tile-based rasterizer.
This approach achieves real-time rendering ($>$100\,FPS) at quality
comparable to or exceeding NeRF, while enabling fast training ($\sim$10\,min).

% ---- Rapid Growth ----
Since its introduction, the 3DGS community has grown explosively.
\todo{Cite statistics: N papers in 2 years.}
Researchers have extended 3DGS to diverse tasks including dynamic
scene reconstruction \cite{wu20244dgs, yang2024deformable},
scene editing, generation, and compression.
However, this rapid growth has outpaced systematic organization,
making it challenging for newcomers to navigate the field
and for researchers to benchmark their contributions.

% ---- Contributions ----
This survey makes the following key contributions:
\begin{enumerate}[leftmargin=*]
    \item A \textbf{three-axis taxonomy} that systematically categorizes
          3DGS research across task domains, module focus, and techniques,
          covering \todo{120+} papers.
    \item The \textbf{Drum Tower Dataset}, a new real-world benchmark
          with Chinese Dong minority drum tower architecture, featuring
          intricate geometry and multi-scale details.
    \item A \textbf{comprehensive benchmark} evaluating \todo{12}
          representative methods across \todo{5} datasets with detailed
          analysis of quality, speed, and memory trade-offs.
    \item An in-depth \textbf{analysis} of current limitations and
          future research directions.
\end{enumerate}

% ---- Organization ----
The remainder of this paper is organized as follows.
Section~\ref{sec:related} reviews related surveys.
Section~\ref{sec:taxonomy} presents our three-axis taxonomy.
Section~\ref{sec:benchmark} describes the Drum Tower Dataset and
benchmark methodology.
Section~\ref{sec:analysis} provides experimental analysis and
discussion.
Section~\ref{sec:conclusion} concludes the paper.
